\begin{thebibliography}{1}

\bibitem{shumailov2021sponge}
I. Shumailov, Y. Zhao, D. Bates, N. Papernot, R. Mullins, and R. Anderson, ``Sponge examples: Energy-latency attacks on neural networks,'' in \emph{2021 IEEE European Symposium on Security and Privacy (EuroS\&P)}, IEEE, 2021, pp. 212--231.

\bibitem{sanchez2024attacking}
G. Sánchez, G. Elbez, and V. Hagenmeyer, ``Attacking Learning-based Models in Smart Grids: Current Challenges and New Frontiers,'' in \emph{Proceedings of the 15th ACM International Conference on Future and Sustainable Energy Systems}, 2024.

\bibitem{cina2022sponge}
A. E. Cinà, et al., ``Energy-latency attacks via sponge poisoning,'' \emph{arXiv preprint arXiv:2203.08147}, 2022.

\bibitem{shapira2023phantom}
A. Shapira, A. Zolfi, L. Demetrio, B. Biggio, and A. Shabtai, ``Phantom Sponges: Exploiting Non-Maximum Suppression to Attack Deep Object Detectors,'' in \emph{Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision (WACV)}, 2023.

\bibitem{telintelo2024skip}
J. te Lintelo, S. Koffas, and S. Picek, ``The SkipSponge Attack: Sponge Weight Poisoning of Deep Neural Networks,'' \emph{arXiv preprint arXiv:2402.06357}, 2024.

\bibitem{wang2023energy}
Z. Wang, Y. Huang, et al., ``Energy-Latency Attacks to On-Device Neural Networks via Sponge Poisoning,'' \emph{arXiv preprint arXiv:2305.03888}, 2023.

\bibitem{attention}
A. Vaswani et al., ``Attention is all you need,'' in \emph{Advances in Neural Information Processing Systems}, 2017, pp. 5998--6008.

\bibitem{steganographic2026}
G. Sánchez, M. Qasim, G. Elbez, and V. Hagenmeyer, ``Steganographic Data Exfiltration for Model Stealing: A Case Study on Energy Critical Infrastructure IEC 61850 Datasets,'' in \emph{IEEE Big Data}, 2026.

\bibitem{hong2024sponge}
J. Hong, et al., ``Sponge Attack Against Multi-Exit Networks with Data Poisoning,'' \emph{IEEE Access}, vol. 12, 2024.

\bibitem{telintelo2025skipsponge}
J. te Lintelo, S. Koffas, and S. Picek, ``The SkipSponge Attack: Sponge Weight Poisoning of Deep Neural Networks,'' \emph{ITU Journal on Future and Evolving Technologies}, vol. 6, no. 3, 2025.

\bibitem{hasan2025sensing}
S. M. Hasan, H. Zangoti, et al., ``Sponge Attacks on Sensing AI: Energy-Latency Vulnerabilities and Defense via Model Pruning,'' \emph{arXiv preprint arXiv:2505.06454}, 2025.

\bibitem{sanchez2025lightweight}
G. Sánchez, G. Elbez, and V. Hagenmeyer, ``Lightweight Moving Target Defense for Robust Intrusion Detection in Smart Grids,'' in \emph{Proceedings of the IEEE International Conference on Communications, Control, and Computing Technologies for Smart Grids (SmartGridComm)}, 2025.

\end{thebibliography}
